Ethnographic Study of Online Entertainers

### 2017-03-16 proposals/Brown_Magic_Grant_2017/main.tex
relatively significant revisions to problem and insight

\definecolor{Blue}{HTML}{0000FF}
\newcommand{\topic}[1]{{\color{Blue}#1}}
% \renewcommand{\topic}[1]{{#1}}
\definecolor{Gray}{HTML}{BABABA}
\newcommand{\ugh}[1]{{\color{Gray}\textit{#1}}}
\newenvironment{blah}{\par\color{Gray}\itshape}{\par}

@@

\section*{Situation}
\topic{In the past decade
online video sharing % \ali{, such as YouTube and Vimeo,}
and streaming sites % \ali{, like Twitch and others,}
have increasingly taken on characteristics of
Deweyan publics~\cite{dewey2012public,disalvo2009design}.}
The ``YouTubers'' and ``streamers'' who use these platforms have come to rely both on
the technical infrastructure these platforms provide \textit{as well as}
the sociality of the venue itself.
These sites have in some sense democratized media and entertainment for online performers by
making the avenues to large audiences substantially more accessible.
In these transformations into publics, these sites
have grown into bustling hubs for artists to engage with audiences and practice their trade.

\section*{Problem}
\topic{Recently, tensions between online performers and the platforms on which they work have shaken people's
assumptions regarding the nature of the space they're using.}
Performers on now--defunct Vine demanded --- unsuccessfully --- to be better compensated~\cite{vineWantsMoney,vineInsiderMeeting};
YouTubers meanwhile struggle to defend their unconventional performance work~\cite{h3h3Lawsuit}
and individually and collectively field accusations of misconduct of various forms~\cite{youtubeDramaResponses}.
\ali{now i want to say something like
``these are all just symptoms of the fact that YouTube and these other platforms aren't the public spaces they were originally perceived to be;
in fact these are carefully controlled spaces, and we need to better understand this stuff''}

\section*{Insight}
\topic{Studying contemporary developments through an established lens can be a helpful way of
making sense of the phenomena we've seen so far, and
provide us with some grounding to make predictions and inform future design decisions.}
We've done this before, studying crowdsourcing and gig work as
an instantiation of \textit{piecework}~\cite{pieceworkCrowdworkGigwork};
this new framing of an existing domain with historical context not only
provided us with guidance toward some of the major threads of research
in the field of crowdsourcing, but
also informed some of the potential futures of crowdsourcing and
the even faster--growing gig work labor markets.

To some extent, we can liken online performance art on YouTube, Twitch, and other platforms to street performances.
Online performers
--- streamers and YouTubers in particular ---
tend not to have the formal training of many conventional performers and artists,
instead growing in proficiency and becoming more skilled principally through their practice;
in much the same way,
street performers are 
%
Online performance artists have the ability
to engage directly with their audiences and tailor their performances to the desires of the crowd
--- YouTubers call upon their audiences to like, dislike, and submit comments to the video, while
    streamers literally change their performances on the fly to suit the whims of those watching and chatting;
meanwhile, the most successful street performers typically engage participants,
drawing them into the performance itself, even bantering with passersby.
%
Finally, online performance has increasingly become a site of political engagement and activism,
sometimes subverting authority
--- in these cases online moderators and algorithms ---
in the process of or for the purpose of making a point;
the history of street performance as a channel for political action and advocacy has been deeply studied
(see, for example, \cite{harrison1990drawing,cohen1998radical}).
\section*{Incomplete!}
\section*{Proposal}

>>>>

\definecolor{Blue}{HTML}{0000FF}
\newcommand{\topic}[1]{{\color{Blue}#1}}
% \renewcommand{\topic}[1]{{#1}}
\definecolor{Gray}{HTML}{A0A0A0}
\newcommand{\ugh}[1]{{\color{Gray}#1}}
\newenvironment{blah}{\par\color{Gray}\itshape}{\par}

@@

\section*{Situation}
\topic{Online video sharing % \ali{, such as YouTube and Vimeo,}
and streaming sites % \ali{, like Twitch and others,}
have garnered the attention of audiences around the world
in the span of little more than a decade.} % taken on characteristics of
So widely used are these platforms
that it would be fair to call them ``publics'' in the Deweyan sense~\cite{dewey2012public,disalvo2009design};
content producers
--- ``YouTubers'' and ``streamers'' ---
have begun to make their livings on these sites.
Indeed, these performers have come to rely on
the technical infrastructure these platforms provide as well as
the access to audiences that these platforms provide as well.
In doing so, these sites have in a real sense democratized media and entertainment industries by
making the avenues to large audiences substantially more accessible.
And thus, in these transformations into publics, what started as
a recreational hobby
has become a bustling hub for professional artists to
engage with audiences,
practice their craft, and 
hone their skills~\cite{Hamilton:2014:STF:2611105.2557048,Zhang:2015:CIL:2736084.2736091}.

\section*{Problem}
\topic{But recently, tensions between these online performers and the platforms on which they work have shaken people's
assumptions regarding the nature of the space they're using.}
Performers on now--defunct Vine demanded --- unsuccessfully --- to be better compensated~\cite{vineWantsMoney,vineInsiderMeeting};
YouTubers meanwhile struggle to defend their unconventional performance work~\cite{h3h3Lawsuit}
and individually and collectively field accusations of misconduct of various forms~\cite{youtubeDramaResponses}.
\ugh{These events fundamentally speak to the contentious politics of an emerging form of expression and art,
and specifically to the uncertainty regarding the status of these artists who tend to subvert not just
the norms of their chosen arts, but also
orthodox labels of performers in the first place.}

\section*{Insight}
\topic{Studying contemporary phenomena through an established framing
can be a helpful way of
making sense of what we've seen so far, as well as
providing us with some grounding to make predictions.}
We've done this before, studying crowd work and gig work as
a modern instantiation of \textit{piecework}~\cite{pieceworkCrowdworkGigwork}.
This positioning of an existing subject in historical context not only
provided us with guidance toward some of the major threads of research
in crowdsourcing, but
also informed some of the potential futures of crowdsourcing and
the ever growing gig work labor markets.

\topic{We can liken some aspects of online performance art on YouTube and Twitch
to street performance and busking,
allowing us to relate ostensibly new phenomena to robust scholarship.}
As is the case with street performance,
online performance isn't restricted to classically trained professionals;
instead, performers (both online and in the streets) practice more experimental 
drawing on their candid interaction with their audiences,
which can change from one performance to the next.

\begin{blah}
Online performers
tend not to have the formal training of many conventional performers and artists,
instead growing in proficiency and becoming more skilled principally through their practice;
in much the same way,
street performers are 
%
Online performance artists have the ability
to engage directly with their audiences and tailor their performances to the desires of the crowd
--- YouTubers call upon their audiences to like, dislike, and submit comments to the video, while
    streamers literally change their performances on the fly to suit the whims of those watching and chatting;
meanwhile, the most successful street performers typically engage participants,
drawing them into the performance itself, even bantering with passersby.
%
Finally, online performance has increasingly become a site of political engagement and activism,
sometimes subverting authority
--- in these cases online moderators and algorithms ---
in the process of or for the purpose of making a point;
the history of street performance as a channel for political action and advocacy has been deeply studied
(see, for example, \cite{harrison1990drawing,cohen1998radical}).
\end{blah}

\section*{Incomplete!}
\section*{Proposal}
